Simulations of Ballot Polling Risk-Limiting Audits

### 2023-02-13 providence/sections/pilot.tex
alpha and section not Section

For this pilot audit, extensive planning of the round schedule was not necessary because the margin was large enough that relatively few ballots were needed to achieve the high probability of stopping. In Section~\ref{sec:workload} we consider a larger state-wide contest in Virginia, where selecting the round schedule has more significant implications. Virginia also currently uses ballot polling RLAs, whereas Rhode Island primarily uses batch comparison RLAs. Some of the ideas introduced in Section~\ref{sec:workload} provide a context for this pilot case as well.

For the sake of analysis, the selection order of the ballots sampled during the pilot was also recorded. Figure~\ref{fig:pilot_sequence} shows the cumulative tally of winner ballots after each new ballot in the selection order is added to the sample. We observe two interesting phenomena in this particular sample's selection order. 
\begin{description}
\item First, an SO \BRAVO audit of this sample stops because the \BRAVO condition is met when the sample (orange line) surpasses the minimum number of winner ballots (blue line) earlier in the sample.\footnote{Such cases also provide insight into how \Providence is a sharper test in expectation because SO \BRAVO ignores information from the rest of the sample after the \BRAVO condition is met at some point earlier in the selection order.} EoR \BRAVO, however, does not stop. It might be difficult to explain to the public why SO \BRAVO stops in more extreme cases like this, where the condition is met early in the sample, but poor evidence for the alternative hypothesis in the rest of the sample is ignored. 
\item Second, only $5$ of the first $11$ ballots were for the announced winner. A first round of size $11$ would have resulted in a smaller average total number of ballots drawn, but would have provided a misleading sample (suggesting that the winner was incorrectly reported) due to a too-small sample size. 
\end{description}
Both these ideas are addressed more thoroughly in Section~\ref{sec:workload}.
%first draft, needed to be shorter: ...stopping condition for the cumulative tally at the end of the round would see that the \BRAVO condition is not met (and would be correct that an EoR \BRAVO audit would require a second round for this sample). On the other hand, an SO \BRAVO audit does stop already in the first round because the cumulative sample met the \BRAVO stopping condition at some point before all the ballots had been added to the sample. In Figure~\ref{fig:pilot_sequence}, this is seen when the observed sample (orange) line touches (or goes above) the minimum winner ballots (blue) line which gives the \BRAVO stopping condition (i.e. a \BRAVO condition is met whenever the number of winner ballots in the sample is at least the number indicated by that blue line). While relatively rare, there also possible cases of selection order where the SO \BRAVO condition is met very early but then bad luck is had thereafter and the ultimate cumulative sample at the end of the round gives very poor evidence that the announced tally is correct. There could even be fewer ballots for the winner than the loser. Such cases might be difficult for an election official to explain to the public, especially when the evidence at the end of the round provides poor support for the alternative hypothesis.\footnote{Such cases also provide insight into how \Providence is a tighter test in expectation because SO \BRAVO ignores information from the rest of the sample after the \BRAVO condition is met at some point earlier in the selection order.} In Section~\ref{sec:workload} we address such cases more thoroughly.

>>>>

For this pilot audit, extensive planning of the round schedule was not necessary because the margin was large enough that relatively few ballots were needed to achieve the high probability of stopping. In section~\ref{sec:workload} we consider a larger state-wide contest in Virginia, where selecting the round schedule has more significant implications. Virginia also currently uses ballot polling RLAs, whereas Rhode Island primarily uses batch comparison RLAs. Some of the ideas introduced in section~\ref{sec:workload} provide a context for this pilot case as well.

For the sake of analysis, the selection order of the ballots sampled during the pilot was also recorded. Figure~\ref{fig:pilot_sequence} shows the cumulative tally of winner ballots after each new ballot in the selection order is added to the sample. We observe two interesting phenomena in this particular sample's selection order. 
\begin{description}
\item First, an SO \BRAVO audit of this sample stops because the \BRAVO condition is met when the sample (orange line) surpasses the minimum number of winner ballots (blue line) earlier in the sample.\footnote{Such cases also provide insight into how \Providence is a sharper test in expectation because SO \BRAVO ignores information from the rest of the sample after the \BRAVO condition is met at some point earlier in the selection order.} EoR \BRAVO, however, does not stop. It might be difficult to explain to the public why SO \BRAVO stops in more extreme cases like this, where the condition is met early in the sample, but poor evidence for the alternative hypothesis in the rest of the sample is ignored. 
\item Second, only $5$ of the first $11$ ballots were for the announced winner. A first round of size $11$ would have resulted in a smaller average total number of ballots drawn, but would have provided a misleading sample (suggesting that the winner was incorrectly reported) due to a too-small sample size. 
\end{description}
Both these ideas are addressed more thoroughly in section~\ref{sec:workload}.
%first draft, needed to be shorter: ...stopping condition for the cumulative tally at the end of the round would see that the \BRAVO condition is not met (and would be correct that an EoR \BRAVO audit would require a second round for this sample). On the other hand, an SO \BRAVO audit does stop already in the first round because the cumulative sample met the \BRAVO stopping condition at some point before all the ballots had been added to the sample. In Figure~\ref{fig:pilot_sequence}, this is seen when the observed sample (orange) line touches (or goes above) the minimum winner ballots (blue) line which gives the \BRAVO stopping condition (i.e. a \BRAVO condition is met whenever the number of winner ballots in the sample is at least the number indicated by that blue line). While relatively rare, there also possible cases of selection order where the SO \BRAVO condition is met very early but then bad luck is had thereafter and the ultimate cumulative sample at the end of the round gives very poor evidence that the announced tally is correct. There could even be fewer ballots for the winner than the loser. Such cases might be difficult for an election official to explain to the public, especially when the evidence at the end of the round provides poor support for the alternative hypothesis.\footnote{Such cases also provide insight into how \Providence is a tighter test in expectation because SO \BRAVO ignores information from the rest of the sample after the \BRAVO condition is met at some point earlier in the selection order.} In Section~\ref{sec:workload} we address such cases more thoroughly.

### 2023-02-13 providence/sections/workload.tex
alpha and section not Section

Importantly, note that \Minerva does not appear in the analysis in this section. 
Questions about the efficiency of \Minerva for its necessarily fixed round schedules are addressed in Section~\ref{sec:sims}, but in this section round sizes are chosen to have specified probabilities of stopping given previous samples. \Minerva is not known to be risk-limiting in this setting, and thus cannot be used for RLAs that proceed in this way.

@@

Recall from Section~\ref{sec:pilot} that the pilot \Providence RLA performed in Providence, Rhode Island had an SO \BRAVO \emph{misleading sequence}. In particular, the audit passed with an SO \BRAVO risk measure of $0.0541$ but the final cumulative tally of the sample gives a \BRAVO risk measure of $0.366$.

>>>>

Importantly, note that \Minerva does not appear in the analysis in this section. 
Questions about the efficiency of \Minerva for its necessarily fixed round schedules are addressed in section~\ref{sec:sims}, but in this section round sizes are chosen to have specified probabilities of stopping given previous samples. \Minerva is not known to be risk-limiting in this setting, and thus cannot be used for RLAs that proceed in this way.

@@

Recall from section~\ref{sec:pilot} that the pilot \Providence RLA performed in Providence, Rhode Island had an SO \BRAVO \emph{misleading sequence}. In particular, the audit passed with an SO \BRAVO risk measure of $0.0541$ but the final cumulative tally of the sample gives a \BRAVO risk measure of $0.366$.
